---
title: "Reproducible Data Analysis"
output:
  html_document:
    toc: false
    theme: united
---

# Reproducible Data Analysis - Peer Assessment

```{r}
require(ggplot2)
require(Hmisc)
```

## Load unzipped csv file from the working directory
```{r}
if(!exists("activity")){
    activity <- read.csv("activity.csv",header=TRUE, sep=",")
}
```

##  Determining mean total number of steps taken per day
```{r}
stepsTakenEveryDay <- tapply(activity$steps, activity$date,sum, na.rm=T)
```

## Histogram of the total number of steps taken each day
```{r echo=FALSE}
qplot(stepsTakenEveryDay, binwidth=1000, xlab="Total number of steps taken each day")
```

## Mean and Median total number of steps taken per day
```{r}
meanStepsTakenEachDay <- mean(stepsTakenEveryDay)
medianStepsTakenEachDay <- median(stepsTakenEveryDay)
```

## Get average number of steps per 5 minutes interval
```{r}
meanStepsPerTimeInterval <- aggregate(steps ~ interval, activity, FUN = mean, na.rm=TRUE)
```

# Time-Series Plot
```{r}
ggplot(data=meanStepsPerTimeInterval, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("Average number of steps taken") 
```

## 5-minute interval, from all the days in the dataset, contains the maximum average number of steps?
```{r}
mostStepsInterval <- which.max(meanStepsPerTimeInterval$steps)
intervalMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", meanStepsPerTimeInterval[mostStepsInterval,'interval'])
```

## Report the total number of missing values in the dataset
numNA <- sum(is.na(activity$steps))

## Creating a new dataset that is equal to the original dataset but with the missing data filled in
## using the "Hmisc" package to "impute" missing values (insert means from the sample)
```{r}
activityFilled <- activity
activityFilled$steps <- impute(activity$steps, fun=mean)
```

## Histogram of the total number of steps taken each day
```{r}
stepsTakenEveryDayFilled <- tapply(activityFilled$steps, activityFilled$date, sum)
qplot(stepsTakenEveryDayFilled, xlab='Total steps per day (with filled missing values)', ylab='Frequency', binwidth=500)
```

## Calculate and report the mean and median total number of steps taken per day
```{r}
meanStepsTakenEachDayFilled <- mean(stepsTakenEveryDayFilled)
meanStepsTakenEachDayFilled <- median(stepsTakenEveryDayFilled)
```

## Creating a new variable with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
getDayType <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("Not a valid date")
}

activityFilled$date <- as.Date(activityFilled$date)
activityFilled$dayType <- sapply(activityFilled$date, FUN=getDayType)
```

## Making a time series plot showing weekday and weekend pattern

```{r}

meanFilledStepsPerTimeInterval <- aggregate(steps ~ interval + dayType, activityFilled, mean)
ggplot(meanFilledStepsPerTimeInterval, aes(interval, steps)) + geom_line() + facet_grid(dayType ~ .) +
    xlab("5-minute interval") + ylab("Number of steps (after filling for missing values)")

```

